Name: Sayed Ebrahim Mustafa Alawi
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
#importing packages to use

library(dplyr)
library(ggplot2)
library(reshape2)
library(gridExtra)
library(forcats)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
#this data is from kaggle 
df = read.csv('googleplaystore.csv')
```

This dataset is from kaggle and it is a dataset about the applications 
available on the google app store.

# Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
#dimensions of the data frame
head(df)
dim(df)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
#Summarise the different Variables so we have an idea of what we're working with
summary(df)
```
In this dataset there are 10841 observation and 13 variables.

###Categories

```{r echo= FALSE,message=FALSE, warning=FALSE}
#plotting a bar graph for categories

df = subset(df, Category != '1.9')


ggplot(aes(x = Category), data = df)+
  geom_bar(fill = 'violetred2')+
  coord_flip()+
  ggtitle("Categories")


```

There are more family applications on google apps than any other category 
followed by health and fitness apps. There are less than 500 applications for
most of the categories on the list.

###Price

```{r echo= FALSE, message=FALSE, warning=FALSE}
df$price = as.numeric(gsub("\\$", "", df$Price))
summary(subset(df, price != 0)$price)

ggplot(aes(x = price), data = subset(df, price != 0))+
  geom_histogram()

ggplot(aes(x = price), data = subset(df, price != 0 & price <= 10))+
  geom_histogram(binwidth = 0.1)
```

Most apps are under 10$ with some prices point being more frequent than others;
(0.99,1.99 etc.)


###Installs

```{r echo= FALSE,message=FALSE, warning=FALSE}
#plotting a bar graph for level of installs.

df = subset(df, df$Installs != 'Free')
ggplot(aes(x = Installs), data = df )+
  geom_bar(fill = 'violetred2')+
  coord_flip()


```

###Rating

```{r echo= FALSE,message=FALSE, warning=FALSE}
#A histogram for the distribution of ratings in the google play dataset
med = median(subset(df$Rating, df$Rating >= 0.01))

ggplot(aes(x = Rating), data = df )+
  geom_histogram(binwidth = 0.1, fill = 'salmon3')+
  xlim(1,5)+ 
  geom_vline(xintercept = med, col = 'red')

summary(df$Rating)
```

The ratings of the apps on the app store follows a left skewed distribution.
most apps are rated above 4.0 and about half are rated between 4.3 and 5.



###Type

```{r echo= FALSE,message=FALSE, warning=FALSE}
#There are two types paid and unpaid
d = subset(df, (Type == 'Free' | Type == 'Paid'))

temp <- d%>%
  group_by(Type)%>%
  summarise(n = n())
  
#pie chart
ggplot(aes(x = '', y = n, fill = Type), data = temp )+
  geom_bar(stat = 'identity')+
  coord_polar('y', start = 0)+
  theme_void()


```

There are a lot more free apps in the app store than paid apps. But we need 
to keep in mind that some of these free apps have in app purchases or they are
free to download but not to use.

####Android Version

```{r echo= FALSE, message=FALSE, warning=FALSE}
#creating a tempory dataset
temp <-df%>%
  group_by(Android.Ver)%>%
  summarise(n = n())

ver_df<-subset(temp, (Android.Ver != 'NaN' & n >10))

ggplot(aes(x = Android.Ver, y = n), data = ver_df)+
geom_bar(stat = 'identity', fill = 'springgreen3')+
coord_flip()

```

The more recent your version of android is the more apps you will be able to
download and use.

###Content Rating

```{r echo= FALSE,message=FALSE, warning=FALSE}
#simple bar plot
ggplot(aes(x = Content.Rating), data = df)+
  geom_bar()+
  coord_flip()+
  scale_y_log10()
```

Most apps are suitable for people under the age of 18, with the apps that have 
an E rating more than all the other categories combined.

###Reviews

```{r echo= FALSE, message=FALSE, warning=FALSE}
#change the reviews column into a column of numbers
df$Reviews = as.numeric(df$Reviews)
ggplot(aes(x = Reviews), data = df)+
  geom_histogram()+
  scale_x_log10()+
  geom_vline(xintercept = median(df$Reviews), color = 'red')
```

###Size

```{r echo= FALSE, message=FALSE, warning=FALSE}
#cleaning up the column to turn it into a form that we can use
dfx = subset(df, Size != 'Varies with device')
#convert all sizes to M and have them be numerical so we can run operations on 
#them
condition = grepl('M', dfx$Size)
if_true = as.numeric(gsub("[a-zA-Z ]", "" , dfx$Size))
if_false = as.numeric(gsub("[a-zA-Z ]", "", dfx$Size))/1024

dfx$Size = ifelse(condition == TRUE, if_true,if_false)

# a simple plot
ggplot(aes(x = round(Size)), data = dfx)+
  geom_histogram(fun.y = count, geom ='line' )+
  geom_vline(xintercept = median(subset(dfx,!is.na(dfx$Size))$Size), col = 'red')+
  geom_vline(xintercept = mean(subset(dfx,!is.na(dfx$Size))$Size), col = 'blue')
  scale_y_log10()
  
  
```

(The blue line is the mean and the red is the median). The average 


###Genres

```{r echo= FALSE,message=FALSE, warning=FALSE}
#Plotting the most popular Genres

#group the genres to find which ones are the most frequent
topgenres = group_by(df, Genres)%>%
  summarise(n = n())%>%
  arrange(desc(n))

#removing the rows that are not the most frequent
#I am going to use this variable again in my bivariate analysis
topgenres = head(topgenres,15)
mask= df$Genres %in% topgenres$Genres
topgenres = df[mask,]

#simple bar plot
ggplot(aes(x = Genres), data = topgenres)+
  geom_bar(fill = 'violetred2')+
  coord_flip()

```

# Univariate Analysis

### What is the structure of your dataset?

>In this dataset there is infortmation about 10841 apps that are available in 
the google app store and 13 columns. 

>About some of the columns:

Type - Free or paid

Categories - There are 33 different Categories (once you remove 1.9)

Installs- There are 21 different levels for installs that tell us how many times
the app has been installed. The levels are not evenly spaced. (once you remove
free)

Rating - The apps can be rated from 1-5

Genres - There are 120 different Genres


Content.Rating - An application can be given 5 different content rating;
Teen, Mature 17+, Everyone 10+, Everyone, Adults only 18+. Some of the 
applications are not rated



>The columns in our dataset - Category, Installs, Type, Content.Rating, 
Genres, current.ver, Android.ver., last.updated, Apps ,Rating, Reviews, size and 
price.

### What is/are the main feature(s) of interest in your dataset?

> Rating is my main feature of interest in this dataset. I would like to see how
the different features affect the rating.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

>I suspect that Genres, Category, Type, and Price affect the rating,

### Did you create any new variables from existing variables in the dataset?
>I have, only to use them temporarily to show a cleaner more relevant plot.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
>


# Bivariate Plots Section

#The relationship between rating and other features.

```{r echo= FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x =Rating, y=Reviews), data = df)+
  geom_jitter(alpha = 0.06)+
  geom_line(stat = 'summary', fun.y = mean)+
  geom_smooth(method = 'lm')
```


```{r echo= FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}

ggplot(aes(x = Rating , y = price ), data = subset(df, Type == 'Paid'))+
  geom_jitter(alpha = 0.3)+
  ylim(0,25)+
  geom_line(stat = 'summary', fun.y = mean)+
  geom_smooth(method = 'lm')

summary(subset(df, Type == 'Paid')$Rating)
```

>Rating statistics for free apps

```{r echo= FALSE, message=FALSE, warning=FALSE}
summary(subset(df, Type == 'Free')$Rating)
```


>Paid apps are slightly better rated than free apps.


```{r echo= FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x =Rating , y =Size ), data = dfx)+
  geom_jitter(alpha = 0.05)+
  geom_smooth(method = 'lm')+
  geom_line(stat = 'summary', fun.y = mean)

cor.test(dfx$Rating,dfx$Size)
```

```{r echo= FALSE, message=FALSE, warning=FALSE}
#plot the average rating for each category
dfcat <- subset(df, !is.na(Rating))
dfcat <- dfcat%>%
  group_by(Category)%>%
  summarise(Rating = mean(as.numeric(Rating)))

#plot
ggplot(aes(x = Category, y = Rating), data = dfcat)+
  geom_bar(stat="identity", position="dodge")+
  coord_flip(ylim = c(3.8,4.5))
  
```

>Dating apps have the lowest average rating and family apps have the most 
average rating


```{r echo= FALSE, message=FALSE, warning=FALSE}
topgenres1 <- subset(topgenres, !is.na(Rating))

topgenres2 <- topgenres1%>%
  group_by(Genres)%>%
  summarise(Rating = mean(as.numeric(Rating)))

ggplot(aes(x = Genres, y = Rating), data = topgenres2)+
  geom_bar(stat = 'identity', position = 'dodge')+
  coord_flip(ylim = c((min(topgenres1$Rating) - 0.1),4.35))

ggplot(aes(x = Genres, y = Rating), data = topgenres1)+
  geom_jitter(alpha = 0.3)+
  coord_flip()

```

>When we look at the most popular genres we find that personalization is on 
averahe the best rated genre followed by education.

```{r echo= FALSE, message=FALSE, warning=FALSE}
ggplot(aes(y = Rating, x = Content.Rating), data = df)+
  geom_jitter(alpha = 0.1, color = 'royalblue1')
```

#Price

##The relationship between price and category

```{r echo= FALSE, message=FALSE, warning=FALSE}
paidapp = subset(df, Type == 'Paid')

ggplot(aes(x =Category , y =price ), data = paidapp)+
  geom_point(alpha = 0.2, color = 'royalblue1')+
  coord_flip(ylim = c(0, quantile(paidapp$price,0.95)))

```


```{r echo= FALSE, message=FALSE, warning=FALSE}
paidappc <-paidapp%>%
  group_by(Category)%>%
  summarise(mean_price = mean(price), n = n(), median_price = median(price))

ggplot(aes(x =Category, y =mean_price ), data = paidappc)+
  geom_bar(stat = 'identity', position = 'dodge', fill = 'royalblue1')+
  coord_flip()+
  geom_text(aes(label = n), hjust=-0.1)

```

The numbers right next to the end of each bar is the number of paid 
applications in that category

When we look at the bar plot above we see that finance applications are the 
most expensive applications on the google app store. I want to explore this 
further I am going to now plot the median price for all the categories and a 
point plot with the price on a logarithmic scale to get a better idea of what
is happening.

```{r echo= FALSE, message=FALSE, warning=FALSE}
 ggplot(aes(y = median_price, x = Category ), data = paidappc)+
  geom_bar(stat = 'identity', fill = 'royalblue1')+
  coord_flip()+
  geom_text(aes(label = n), hjust = -0.1)

ggplot(aes(y = price, x = Category), data = paidapp)+
  geom_point(alpha = 0.2)+
  coord_flip()+
  scale_y_log10()

temp <- head(subset(df, Category == 'FINANCE' & Type == 'Paid'))%>%
  select(App, Category, price, Genres)

temp[order(temp$price, decreasing = TRUE),]%>%
  head(5)
```

I found that this is due to the apps that were meant to show off wealth if 
we were to not count these apps as part of our statistic; if we were to assume
that they do not have a practical purpose and were not meant to be taken 
seriously this will lead us to the conclusion that the mean/median price is  
a lot lower.


```{r echo= FALSE, message=FALSE, warning=FALSE}
temp <- subset(df, Category == 'LIFESTYLE' & Type == 'Paid')%>%
  select(App, Category, price, Genres)

temp[order(temp$price, decreasing = TRUE),]%>%
  head(5)

```

In the lifestyle category we see that these same sort of applications is 
distorting our statistic.

```{r echo= FALSE, message=FALSE, warning=FALSE}
subset(df, Category == 'EVENTS' & Type == 'Paid')%>%
  select(App,Category,price, Genres)

```

In the events category there is only one paid app and
that app has only been installed between 1-5 times. If we were to visually
inspect the point plot we would see that on average the medical apps are 
actually more expensive on average.

##The relationship between Genres and Category

let us look at the most frequent Genres and see what Categories makes up each 
genre

```{r echo= FALSE, message=FALSE, warning=FALSE}
temp <- df[df$Category %in% topgenres$Category,]

ggplot(aes(x = Category), data = df)+
  geom_bar(aes(fill = Genres))+
  coord_flip()+
  theme(legend.position="none")

```
 
 Here we see that the family and game categories are made up of multiple genres.
 Let us explore this further.
 
```{r echo= FALSE, message=FALSE, warning=FALSE}
temp <- subset(df, Category == 'GAME')%>%
  group_by(Genres)%>%
  summarise(n = n())

ggplot(aes( x = Genres, y = n), data = temp)+
  geom_bar(stat = 'identity', fill = 'tomato')+
  coord_flip()+
  geom_text(aes(label = n), hjust = -0.1)
```
 
 The most popular Genre in the game Category is action with 365 apps followed
 by arcade with 220 apps and racing with 98 apps.
 
```{r echo= FALSE, message=FALSE, warning=FALSE}
temp <- subset(df, Category == 'FAMILY')%>%
  group_by(Genres)%>%
  summarise(n = n())%>%
  arrange()

temp<- temp[order(temp$n, decreasing = TRUE),]

temp <- head(temp, 20)

ggplot(aes( x = Genres, y = n), data = temp)+
  geom_bar(stat = 'identity', fill = 'tomato')+
  coord_flip()+
  geom_text(aes(label = n), hjust = -0.1)
```
 
 Above are the top 20 genres based on frequency that make up the family 
 category. The most frequent apps in the family category are entertainment apps 
 (487) followed by education apps (420) and simulation apps(189).

```{r echo= FALSE, message=FALSE, warning=FALSE}
ggplot(aes (x = Genres), data = topgenres)+
  geom_bar(aes(fill = Category))+
  coord_flip()
```

Here we see that a genre can fall under different categories, for example
most of the apps in the entertainment genre are in the family category although
there is an entertainment category. We see the same thing with the education
genre; most of the apps in the education genre are under the family category
although there is a category for educational apps.

##The relationship between the most frequent Genres and price

```{r echo= FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x= Genres, y = price), data = subset(topgenres, Type =='Paid'))+
  geom_jitter(stat = 'identity', alpha = 0.5, color = 'slateblue1')+
  coord_flip(ylim = c(0,50))
```

Here we see that the most expensive apps are in the medical genre we also see 
that most of the tools are under 10$.

##The relationship between price and other features

```{r echo= FALSE, message=FALSE, warning=FALSE}
#simple plot of price*size
ggplot(aes(x = price, y = Size), data = subset(dfx, Type == 'Paid'))+
  geom_jitter(alpha = 0.1)+
  xlim(0,100)+
  geom_smooth(method = 'lm')

#temporary dataset where the type is paid and without the 'troll' apps
temp <- subset(dfx, Type == 'Paid' & price <300)

#correlation
cor.test(temp$price, temp$Size)
```

```{r echo= FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = Reviews, y = price), data = subset(df, Type == 'Paid'))+
  geom_point()+
  geom_smooth()+
  scale_y_log10()

cor.test(df$Reviews, df$price)
```

##The relationship between Installs and other Feautures

```{r echo= FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = Installs, y = price), data = subset(df, Type == 'Paid') )+
  geom_jitter(alpha = 0.1)+
  coord_flip()

```

The cheaper the app the more likely it will be downloaded and installed.

```{r echo= FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = Installs), data = df)+
  geom_bar(aes(fill = Category))+
  coord_flip()+
  theme(legend.position="none")

```

It does not seem as though the category affects the number of times an app
would be installed, I came to this conclusion since it seems that the different
categories are in pretty much all the 'Installs' levels.

```{r echo= FALSE, message=FALSE, warning=FALSE}

ggplot(aes( x= Installs), data = df)+
  geom_bar(aes(fill = Type))+
  coord_flip()

```
No paid app has been installed more than or equal to 50 million times.

```{r echo= FALSE, message=FALSE, warning=FALSE}
temp <- df%>%
  group_by(Installs)%>%
  summarise(mean = mean(Reviews), median = median(Reviews), max = max(Reviews),
            n = n())

ggplot(aes(x=Installs, y = Reviews), data = df)+
  geom_jitter(alpha = 0.05)+
  coord_flip()

ggplot(aes(x = Installs, y = mean), data = temp)+
  geom_bar(stat = 'identity')+
  coord_flip()+
  geom_text(aes(label = n), hjust = -0.1)

```

When we look at the 5+,10+,50+,100+,500+ levels in the point plot abover there 
are a certain number of reviews that are more frequent. I then created a bar
plot to see the mean number of reviews for each level. 500+ which I assume to be
a level assigned to all applications that are installed between 500 times up until
the next level have more reviews than the times they have been installed.

##The relationship between category and content rating
```{r echo= FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = Category), data = df)+
  geom_bar(aes(fill = Content.Rating))+
  coord_flip()+
  scale_y_log10()

```

The only apps that have a conent rating of Adult only are in the comics 
category. The Dating Category has the most apps with acontent rating of mature.
Every category has more than one application that has a content rating of 
Everyone.


# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!