Name: Sayed Ebrahim Mustafa Alawi
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
#importing packages to use
library(dplyr)
library(ggplot2)
library(reshape2)
library(gridExtra)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
#This dataset is from kaggle 
df = read.csv('googleplaystore.csv')
```

This dataset is from kaggle and it is a dataset about the applications 
available on the google app store.

# Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
#first 5 rows of the dataframe
head(df)
#dimensions of the data frame
dim(df)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
#Summarise the different Variables so we have an idea of what we're working with
summary(df)
```
In this dataset there are 10841 observation and 13 variables.

###Category

```{r echo= FALSE,message=FALSE, warning=FALSE}
#removing the 1.9 category
df = subset(df, Category != '1.9')

#plotting a bar graph for categories
ggplot(aes(x = Category), data = df)+
  geom_bar(fill = 'royalblue2')+
  coord_flip()+
  ggtitle("Categories")

```

There are more applications in the Family Category than any other Category 
followed by the game and tools categories. There are less than 500 applications 
for most of the categories on the list.

###Price

```{r echo= FALSE, message=FALSE, warning=FALSE}
#clean the price feature so I can use it as a float
df$price = as.numeric(gsub("\\$", "", df$Price))

ggplot(aes(x = price), data = df)+
  geom_histogram(fill = 'royalblue2', binwidth = 3)+
  scale_y_log10()+
  ggtitle('Price')
```

Most apps do not cost a thing and we find that there are outliers around the 
400 dollar mark.

###Installs

```{r echo= FALSE,message=FALSE, warning=FALSE}
#Cleaning the installs column
df = subset(df, df$Installs != 'Free')

#plotting a bar graph for level of installs.
ggplot(aes(x = Installs), data = df )+
  geom_bar(fill = 'violetred2')+
  coord_flip()+
  ggtitle('Installs')
```

The level with the most number of installs is 1000,000+, which is 1-5 million. 
The install levels range increases exponentially.

###Rating

```{r echo= FALSE,message=FALSE, warning=FALSE}
#A histogram for the distribution of ratings in the google play dataset
med = median(subset(df$Rating, df$Rating >= 0.01))

#Histogram
ggplot(aes(x = Rating), data = df )+
  geom_histogram(binwidth = 0.1, fill = 'violetred2')+
  xlim(1,5)+ 
  geom_vline(xintercept = med, col = 'blue')+
  ggtitle('Rating')

summary(df$Rating)
```

The ratings of the apps in the app store follows a left skewed distribution.
most apps are rated above 4.0 and about half are rated between 4.3 and 5.
(blue line is the median)


###Type

```{r echo= FALSE,message=FALSE, warning=FALSE}
#There are two types paid and unpaid
df_type = subset(df, (Type == 'Free' | Type == 'Paid'))

temp <- df_type%>%
  group_by(Type)%>%
  summarise(n = n())

#pie chart
ggplot(aes(x = '', y = n, fill = Type), data = temp )+
  geom_bar(stat = 'identity')+
  coord_polar('y', start = 0)+
  theme_void()+
  ggtitle('Type')


```

There are a lot more free apps in the app store than paid apps. But we need 
to keep in mind that some of these free apps have in app purchases.

###Android Version

```{r echo= FALSE, message=FALSE, warning=FALSE}
#creating a tempory dataset
temp <-df%>%
  group_by(Android.Ver)%>%
  summarise(n = n())


ver_df<-subset(temp, (Android.Ver != 'NaN' & n >10))

#bar plot
ggplot(aes(x = Android.Ver, y = n), data = ver_df)+
geom_bar(stat = 'identity', fill = 'royalblue2')+
coord_flip()+
ylab('count')+
xlab('Version Of Android')+
ggtitle('Android Version')

```

The more recent your version of android is the more apps you will be able to
download and use. Some applications require different version of android to 
install depending on your device.

###Content Rating

```{r echo= FALSE,message=FALSE, warning=FALSE}

#simple bar plot
ggplot(aes(x = Content.Rating), data = df)+
  geom_bar(fill = 'royalblue2')+
  coord_flip()+
  scale_y_log10()+
  ggtitle('Content Rating')
```

Most apps are suitable for people under the age of 18. The number of 
applications that are rated for everybody are more than all the other content 
rating combined.

###Reviews

```{r echo= FALSE, message=FALSE, warning=FALSE}
#change the reviews column into a column of numbers
df$Reviews = as.numeric(df$Reviews)

#Histogram
ggplot(aes(x = Reviews), data = df)+
  geom_histogram(fill = 'violetred2')+
  scale_x_log10()+
  geom_vline(xintercept = median(df$Reviews), color = 'blue')+
  ggtitle('Reviews')

#summary
summary(df$Reviews)
```

The numbner of reviews follows a left skewed distribution. About half the apps 
have more than 2747 reviews.(blue line is the median)

###Size

```{r echo= FALSE, message=FALSE, warning=FALSE}
#cleaning up the column to turn it into a form that we can use
df_size = subset(df, Size != 'Varies with device')

#convert all sizes to M and have them be numerical so we can run operations on 
#them
condition = grepl('M', df_size$Size)
if_true = as.numeric(gsub("[a-zA-Z ]", "" , df_size$Size))
if_false = as.numeric(gsub("[a-zA-Z ]", "", df_size$Size))/1024

df_size$Size = ifelse(condition == TRUE, if_true,if_false)

# a simple plot
ggplot(aes(x = round(Size)), data = df_size)+
  geom_histogram(fun.y = count, geom ='line', fill = 'violetred2')+
  geom_vline(xintercept = median(subset(df_size,!is.na(df_size$Size))$Size), col = 'red')+
  geom_vline(xintercept = mean(subset(df_size,!is.na(df_size$Size))$Size), col = 'blue')+
  ggtitle('Size')+
  xlab('Size')
  
  
```

The size of the applications follows a right skewed distribution.
(The blue line is the mean and the red is the median). 


###Genres

```{r echo= FALSE,message=FALSE, warning=FALSE}
#Plotting the most popular Genres

#group the genres to find which ones are the most frequent
topgenres = group_by(df, Genres)%>%
  summarise(n = n())%>%
  arrange(desc(n))

#removing the rows that are not the most frequent
#I am going to use this variable again in my bivariate analysis
topgenres = head(topgenres,15)
mask= df$Genres %in% topgenres$Genres
topgenres = df[mask,]

#simple bar plot
ggplot(aes(x = Genres), data = topgenres)+
  geom_bar(fill = 'violetred2')+
  coord_flip()+
  ggtitle('Genres')

```

# Univariate Analysis

### What is the structure of your dataset?

In this dataset there is infortmation about 10841 apps that are available in 
the google app store and 13 features. 

About some of the columns:

Type - Free or paid
Categories - There are 33 different Categories (once you remove 1.9)

Installs- There are 21 different levels for installs that tell us how many times
the app has been installed. 

Rating - The apps can be rated from 1-5

Genres - There are 120 different Genres

Content.Rating - An application can be given 5 different content ratings;
Teen, Mature (17+), Everyone (10+), Everyone, Adults only (18+). Some of the 
applications are not rated


The different features in our dataset - Category, Installs, Type, 
Content.Rating, Genres, current.ver, Android.ver., last.updated, Apps ,Rating,
Reviews, size and price.

### What is/are the main feature(s) of interest in your dataset?

Rating and price are my main features of interest in this dataset.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

I would like to see if the Genre, Category, Reviews, Rating and type affect the 
price and rating.

### Did you create any new variables from existing variables in the dataset?
I have.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
I changed the price feature from a column of strings to a column of floats.
I changed the size feature in the same way and manipulated it in a way that 
all the values share the same unit.


# Bivariate Plots Section

##The relationship between rating and other features.

###Rating vs. Reviews
```{r echo= FALSE, message=FALSE, warning=FALSE}
#plot of rating vs reviews
ggplot(aes(y =Rating, x=Reviews), data = df)+
  geom_jitter(alpha = 0.06, color = 'royalblue2')+
  geom_smooth(method = 'lm')+
  ggtitle('Rating vs. Reviews')
```

The variability of the rating of an application is not explained by the number 
of reviews.

###Rating vs. Installs

```{r echo= FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x =Installs , y =Rating ), data = df)+
  geom_boxplot(fill = 'violetred2')+
  coord_flip()+
  ggtitle('Rating vs. Installs')

```

We can see that the median rating will be above 4 no matter how many times an
app was installed.

###Rating vs. Price
```{r echo= FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
#plot of rating vs. price
ggplot(aes(x = Rating , y = price), data = df)+
  geom_jitter(alpha = 0.3, color = 'royalblue1')+
  ylim(0,25)+
  geom_line(stat = 'summary', fun.y = mean)+
  ggtitle('Rating vs. Price')
```


Here we see that the mean price is very close to 0 this is because most of the
apps are free. To get a better understanding of the relationship between rating
and price we need to filter by type we're going to do this in the multivariate
section with type as the third variable.


###Rating vs. Type
```{r echo= FALSE, message=FALSE, warning=FALSE}
#plot for rating vs. type
ggplot(aes(x = Type, y = Rating), data = df_type )+
  geom_boxplot(fill = 'violetred2' )+
  ggtitle('Rating vs. Type')
#summary for the paid
summary(subset(df, Type == 'Paid')$Rating)
#summary
summary(subset(df, Type == 'Free')$Rating)
```

Paid apps are slightly better rated than free apps. 

###Rating vs. Size
```{r echo= FALSE, message=FALSE, warning=FALSE}
# Use the df_size dataframe we created in the univariate analysis section
#to plot the relationship between rating and size
ggplot(aes(x =Rating , y =Size ), data = df_size)+
  geom_jitter(alpha = 0.05, color ='royalblue1')+
  geom_smooth(method = 'lm')+
  geom_line(stat = 'summary', fun.y = mean  )+
  ggtitle('Rating vs. Size')

#correlation between rating and size
cor.test(df_size$Rating,df_size$Size)
```

The size of an application does not affect the rating. Users are not bothered
by large app sizes.

### Rating vs. Category

```{r echo= FALSE, message=FALSE, warning=FALSE}
#plot the average rating for each category
dfcat <- subset(df, !is.na(Rating))
dfcat <- dfcat%>%
  group_by(Category)%>%
  summarise(Rating = mean(as.numeric(Rating)))

#plot
ggplot(aes(x = Category, y = Rating), data = dfcat)+
  geom_bar(stat="identity", fill = 'violetred2')+
  coord_flip(ylim = c(3.8,4.5))

```

Dating applications have the lowest average rating while family applications 
have the most average rating.

###Rating vs. Top Genres
```{r echo= FALSE, message=FALSE, warning=FALSE}
#remove all the rows that have no rating
topgenres <- subset(topgenres, !is.na(Rating))

#group the topgenres coulmn
topgenres_group <- topgenres%>%
  group_by(Genres)%>%
  summarise(Rating = mean(as.numeric(Rating)))

#barplot of the mean rating for the most frequent genres
ggplot(aes(x = Genres, y = Rating), data = topgenres_group)+
  geom_bar(stat = 'identity', position = 'dodge', fill = 'royalblue2')+
  coord_flip(ylim = c((min(topgenres_group$Rating) - 0.1),4.35))+
  ggtitle('Rating vs. Top Genres')

#point plot
ggplot(aes(x = Genres, y = Rating), data = topgenres)+
  geom_jitter(alpha = 0.3, color = 'royalblue2')+
  coord_flip()+
  ggtitle('Rating vs. Top Genres')

```

When we look at the most popular genres we find that personalization is on 
average the best rated genre followed by education.

###Rating vs. Content Rating
```{r echo= FALSE, message=FALSE, warning=FALSE}
#creating a box plot without the unrated level in content rating
ggplot(aes(y = Rating, x = Content.Rating),
       data = subset(df, Content.Rating != 'Unrated'))+
  geom_boxplot( fill = 'royalblue1')+
  ggtitle('Rating vs. Content Rating')
```

Applications with a content rating of mature (17+) have a lower rating.



##The relationship between Genres and Category

let us look at the most frequent Genres and see what Categories makes up each 
genre

```{r echo= FALSE, message=FALSE, warning=FALSE}
#plot of the different categories with different colors as the genres
ggplot(aes(x = Category), data = df)+
  geom_bar(aes(fill = Genres))+
  coord_flip()+
  theme(legend.position="none")+
  ggtitle('Genre and Category')
```
 
 Here we see that the family and game categories are made up of multiple genres.
 Let us explore this further.
 
```{r echo= FALSE, message=FALSE, warning=FALSE}
#temporary dataframe that will only be used one and has the number of 
#applications in each genre that make up the Game Category
temp <- subset(df, Category == 'GAME')%>%
  group_by(Genres)%>%
  summarise(n = n())

#plot
ggplot(aes( x = Genres, y = n), data = temp)+
  geom_bar(stat = 'identity', fill = 'tomato')+
  coord_flip()+
  geom_text(aes(label = n), hjust = -0.1)+
  ggtitle("Genre in Game Category")+
  ylab('count')
```
 
 The most popular Genre in the game Category is action with 365 apps followed
 by arcade with 220 apps and racing with 98 apps.
 
```{r echo= FALSE, message=FALSE, warning=FALSE}
#temporary dataframe that will only be used one and has the number of 
#applications in each genre that make up theFamily Category

temp <- subset(df, Category == 'FAMILY')%>%
  group_by(Genres)%>%
  summarise(n = n())%>%
  arrange()

#The top 20 Genres
temp<- temp[order(temp$n, decreasing = TRUE),]
temp <- head(temp, 20)

#plot
ggplot(aes( x = Genres, y = n), data = temp)+
  geom_bar(stat = 'identity', fill = 'tomato')+
  coord_flip()+
  geom_text(aes(label = n), hjust = -0.1)+
  ggtitle("Genre in Family Category")+
  ylab('count')
```
 
 Above are the top 20 genres based on frequency that make up the family 
 category. The most frequent apps in the family category are entertainment apps 
 (487) followed by education apps (420) and simulation apps(189).

```{r echo= FALSE, message=FALSE, warning=FALSE}
#plot of the topgenres and category
ggplot(aes (x = Genres), data = topgenres)+
  geom_bar(aes(fill = Category))+
  coord_flip()+
  ggtitle('Top genres and Category')
```

Here we see that a genre can fall under different categories, for example
most of the apps in the entertainment genre are in the family category although
there is an entertainment category. We see the same thing with the education
genre; most of the apps in the education genre are under the family category
although there is a category for educational apps.


```{r echo= FALSE, message=FALSE, warning=FALSE}
#plot of type by category
ggplot(aes(x= Category), data = df)+
  geom_bar(aes(fill = Type))+
  coord_flip()+
  ggtitle('Category vs. Type')

```

##The relationship between Installs and other Feautures

```{r echo= FALSE, message=FALSE, warning=FALSE}
#plot of install by category
ggplot(aes(x = Installs), data = df)+
  geom_bar(aes(fill = Category))+
  coord_flip()+
  theme(legend.position="none")+
  ggtitle('Installs by Category')

```

It does not seem as though the category affects the number of times an app
would be installed, I came to this conclusion since it seems that the different
categories are in pretty much all the 'Installs' levels.

```{r echo= FALSE, message=FALSE, warning=FALSE}
#Installs vs Type
ggplot(aes( x= Installs), data = df)+
  geom_bar(aes(fill = Type))+
  coord_flip()+
  ggtitle('Installs vs. Type')

```
No paid app has been installed more than or equal to 50 million times.

```{r echo= FALSE, message=FALSE, warning=FALSE}
#Temporary dataset where we group installs and get summaries
temp <- df%>%
  group_by(Installs)%>%
  summarise(mean = mean(Reviews), median = median(Reviews), max = max(Reviews),
            n = n())

#Install vs. Reviews
ggplot(aes(x=Installs, y = Reviews), data = df)+
  geom_jitter(alpha = 0.05, color = 'violetred2')+
  coord_flip()+
  ggtitle('Install vs. Reviews')

#install vs. mean reviews
ggplot(aes(x = Installs, y = mean), data = temp)+
  geom_bar(stat = 'identity', fill = 'royalblue2')+
  coord_flip()+
  geom_text(aes(label = n), hjust = -0.1)+
  ggtitle('Installs vs mean no. of Reviews')+
  ylab('mean reviews')

```

When we look at the 5+,10+,50+,100+,500+ levels in the point plot abover there 
are a certain number of reviews that are more frequent. I then created a bar
plot to see the mean number of reviews for each level. 500+ which I assume to be
a level assigned to all applications that are installed between 500 times up 
until the next level have more reviews than the times they have been installed.

##The relationship between category and content rating
```{r echo= FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = Category), data = df)+
  geom_bar(aes(fill = Content.Rating))+
  coord_flip()+
  scale_y_log10()+
  ggtitle('category vs content rating')

```

The only apps that have a conent rating of Adult only are in the comics 
category. The Dating Category has the most apps with acontent rating of mature.
Every category has more than one application that has a content rating of 
Everyone.


# Bivariate Analysis

In my bivariate analysis I found that the rating of an application is not 
strongly coorelated to any feature in the dataset. I am going to have to work
on the price in the multivarite section in order to inclue type.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

-Paid apps are better rated than free apps.
-

### Did you observe any interesting relationships between the other features \

The Family and games categories have different genres.

### What was the strongest relationship you found?

All the relationships between quantitative variables were pretty weak. (<0.3)

# Multivariate Plots Section

##The relationship between price and Type with other Features

```{r echo= FALSE, message=FALSE, warning=FALSE}

paidapp = subset(df, Type == 'Paid')

ggplot(aes(x =Category , y =price ), data = df)+
  geom_point(alpha = 0.2, color = 'royalblue1')+
  coord_flip(ylim = c(0, quantile(paidapp$price,0.95)))

```


```{r echo= FALSE, message=FALSE, warning=FALSE}
paidappc <-paidapp%>%
  group_by(Category)%>%
  summarise(mean_price = mean(price), n = n(), median_price = median(price))

ggplot(aes(x =Category, y =mean_price ), data = paidappc)+
  geom_bar(stat = 'identity', position = 'dodge', fill = 'royalblue1')+
  coord_flip()+
  geom_text(aes(label = n), hjust=-0.1)

```

The numbers right next to the end of each bar is the number of paid 
applications in that category

When we look at the bar plot above we see that finance applications are the 
most expensive applications on the google app store. I want to explore this 
further I am going to now plot the median price for all the categories and a 
point plot with the price on a logarithmic scale to get a better idea of what
is happening.

```{r echo= FALSE, message=FALSE, warning=FALSE}
 ggplot(aes(y = median_price, x = Category ), data = paidappc)+
  geom_bar(stat = 'identity', fill = 'royalblue1')+
  coord_flip()+
  geom_text(aes(label = n), hjust = -0.1)

ggplot(aes(y = price, x = Category), data = paidapp)+
  geom_point(alpha = 0.2)+
  coord_flip()+
  scale_y_log10()

temp <- head(subset(df, Category == 'FINANCE' & Type == 'Paid'))%>%
  select(App, Category, price, Genres)

temp[order(temp$price, decreasing = TRUE),]%>%
  head(5)
```

I found that this is due to the apps that were meant to show off wealth if 
we were to not count these apps as part of our statistic; if we were to assume
that they do not have a practical purpose and were not meant to be taken 
seriously this will lead us to the conclusion that the mean/median price is  
a lot lower.


```{r echo= FALSE, message=FALSE, warning=FALSE}
temp <- subset(df, Category == 'LIFESTYLE' & Type == 'Paid')%>%
  select(App, Category, price, Genres)

temp[order(temp$price, decreasing = TRUE),]%>%
  head(5)

```

In the lifestyle category we see that these same sort of applications is 
distorting our statistic.

```{r echo= FALSE, message=FALSE, warning=FALSE}
subset(df, Category == 'EVENTS' & Type == 'Paid')%>%
  select(App,Category,price, Genres)

```

In the events category there is only one paid app and
that app has only been installed between 1-5 times. If we were to visually
inspect the point plot we would see that on average the medical apps are 
actually more expensive on average.


```{r echo= FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = Installs, y = price), data = subset(df, Type == 'Paid') )+
  geom_jitter(alpha = 0.1)+
  coord_flip()+
  ggtitle('Installs vs. Type')

```


The cheaper the app the more likely it will be downloaded and installed.


```{r echo=FALSE, Multivariate_Plots}

ggplot(aes(x = Reviews, y = price), data = subset(df, Type = 'Paid'))+
  geom_jitter(aes(color = Android.Ver))+
  theme(legend.position = 'None')+
  ylim(0,50)

```


##The relationship between price and other features

```{r echo= FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x= Genres, y = price), data = subset(topgenres, Type =='Paid'))+
  geom_jitter(stat = 'identity', alpha = 0.5, aes(color = Reviews))+
  coord_flip(ylim = c(0,50))
```

Here we see that the most expensive apps are in the medical genre we also see 
that most of the tools are under 10$.


```{r echo= FALSE, message=FALSE, warning=FALSE}
#simple plot of price*size
ggplot(aes(x = price, y = Size), data = subset(df_size, Type == 'Paid'))+
  geom_jitter(alpha = 0.1)+
  xlim(0,100)+
  geom_smooth(method = 'lm')

#temporary dataset where the type is paid and without the 'troll' apps
temp <- subset(df_size, Type == 'Paid' & price <300)

#correlation
cor.test(temp$price, temp$Size)
```

```{r echo= FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = Reviews, y = price), data = subset(df, Type == 'Paid'))+
  geom_point()+
  geom_smooth()+
  scale_y_log10()

cor.test(df$Reviews, df$price)
```

# Multivariate Analysis


### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!