Name: Sayed Ebrahim Mustafa Alawi
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
#importing packages to use

library(dplyr)
library(ggplot2)
library(reshape2)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
#this data is from kaggle 
df = read.csv('googleplaystore.csv')
```

>This dataset is from kaggle and it is a dataset about the applications 
available on the google app store.

# Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
#dimensions of the data frame
head(df)
dim(df)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
#Summarise the different Variables so we have an idea of what we're working with
summary(df)
```
>In this dataset there are 10841 observation and 13 variables.

###Categories

```{r echo= FALSE,message=FALSE, warning=FALSE}
#plotting a bar graph for categories

df = subset(df, Category != '1.9')


ggplot(aes(x = Category), data = df)+
  geom_bar()+
  coord_flip()+
  ggtitle("Categories")


```

> There are more family applications on google apps then any other category 
followed by health and fitness apps. There are less than 500 application for
most of the categories on the list.

###Installs

```{r echo= FALSE,message=FALSE, warning=FALSE}
#plotting a bar graph for level of installs.

df = subset(df, df$Installs != 'Free')
ggplot(aes(x = Installs), data = df )+
  geom_bar()+
  coord_flip()


```

###Rating

```{r echo= FALSE,message=FALSE, warning=FALSE}
#A histogram for the distribution of ratings in the google play dataset
med = median(subset(df$Rating, df$Rating >= 0.01))

ggplot(aes(x = Rating), data = df )+
  geom_histogram(binwidth = 0.1)+
  xlim(1,5)+ 
  geom_vline(xintercept = med, col = 'red')

summary(df$Rating)
```

>The ratings of the apps on the app store follows a left skewed distribution.
most apps are rated above 4.0 and about half are rated between 4.3 and 5.

###Genres

```{r echo= FALSE,message=FALSE, warning=FALSE}
#Plotting the most popular Genres

#group the genres to find which ones are the most popular 
topgenres = group_by(df, Genres)%>%
  summarise(n = n())%>%
  arrange(desc(n))

#removing the rows that are not the most popular category
#I am going to use this variable again in my bivariate analysis
topgenres = head(topgenres,15)
mask= df$Genres %in% topgenres$Genres
topgenres = df[mask,]

#simple bar plot
ggplot(aes(x = Genres), data = topgenres)+
  geom_bar()+
  coord_flip()

```

###Type

```{r echo= FALSE,message=FALSE, warning=FALSE}
#There are two types paid and unpaid
d = subset(df, (Type == 'Free' | Type == 'Paid'))

ggplot(aes(x = Type), data = d )+
  geom_bar()
```

>There are a lot more free apps in the app store than paid apps. But we need 
to keep in mind that some of these free apps have in app purchases or they are
free to download but not to use.

###Content Rating

```{r echo= FALSE,message=FALSE, warning=FALSE}
#simple bar plot
ggplot(aes(x = Content.Rating), data = df)+
  geom_bar()+
  coord_flip()+
  scale_y_log10()
```

>Most apps are suitable for people under the age of 18, with the apps that have 
an E rating more than all the other categories combined.

###Reviews

```{r echo= FALSE, message=FALSE, warning=FALSE}
#change the reviews column into a column of numbers
df$Reviews = as.numeric(df$Reviews)
ggplot(aes(x = Reviews), data = df)+
  geom_histogram()+
  scale_x_log10()+
  geom_vline(xintercept = median(df$Reviews), color = 'red')
```

###Size

```{r echo= FALSE, message=FALSE, warning=FALSE}
#cleaning up the column to turn it into a form that we can use
dfx = subset(df, Size != 'Varies with device')
#convert all sizes to M and have them be numerical so we can run operations on 
#them
condition = grepl('M', dfx$Size)
if_true = as.numeric(gsub("[a-zA-Z ]", "" , dfx$Size))
if_false = as.numeric(gsub("[a-zA-Z ]", "", dfx$Size))/1024

dfx$Size = ifelse(condition == TRUE, if_true,if_false)

# a simple plot
ggplot(aes(x = round(Size)), data = dfx)+
  geom_histogram(fun.y = count, geom ='line' )+
  geom_vline(xintercept = median(subset(dfx,!is.na(dfx$Size))$Size), col = 'red')+
  geom_vline(xintercept = mean(subset(dfx,!is.na(dfx$Size))$Size), col = 'blue')
  scale_y_log10()
  
  summary(dfx$Size)
```

> (The blue line is the mean and the red is the median). The average 

###Price

```{r echo= FALSE, message=FALSE, warning=FALSE}
df$price = as.numeric(gsub("\\$", "", df$Price))
summary(subset(df, price != 0)$price)

ggplot(aes(x = price), data = subset(df, price != 0))+
  geom_histogram()

ggplot(aes(x = price), data = subset(df, price != 0 & price <= 10))+
  geom_histogram(binwidth = 0.1)
```

>Most apps are under 10$ with some prices point being more frequent than others;
(0.99,1.99 etc.)


# Univariate Analysis

### What is the structure of your dataset?

>In this dataset there is infortmation about 10841 apps that are available in 
the google app store and 13 columns. 

>About some of the columns:

>Type - Free or paid

>Categories - There are 33 different Categories (once you remove 1.9)

>Installs- There are 21 different levels for installs that tell us how many times
>the app has been installed. The levels are not evenly spaced. (once you remove
free)

>Rating - The apps can be rated from 1-5

>Genres - There are 120 different Genres


>Content.Rating - An application can be given 5 different content rating;
Teen, Mature 17+, Everyone 10+, Everyone, Adults only 18+. Some of the 
applications are not rated



>The columns in our dataset - Category, Installs, Type, Content.Rating, 
Genres, current.ver, Android.ver., last.updated, Apps ,Rating, Reviews, size and 
price.

### What is/are the main feature(s) of interest in your dataset?

> Rating is my main feature of interest in this dataset. I would like to see how
the different features affect the rating.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

>I suspect that Genres, Category, Type, and Price affect the rating,

### Did you create any new variables from existing variables in the dataset?
>I have, only to use them temporarily to show a cleaner more relevant plot.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
>


# Bivariate Plots Section

#Rating

```{r echo=FALSE, Bivariate_Plots}

ggplot(aes(x = Rating , y = price ), data = subset(df, Type == 'Paid'))+
  geom_point()+
  ylim(0,50)

summary(subset(df, Type == 'Paid')$Rating)
```

>Rating statistics for free apps

```{r}
summary(subset(df, Type == 'Free')$Rating)
```


>Paid apps are slightly better rated than free apps.


```{r}

ggplot(aes(x =Rating , y =Size ), data = dfx)+
  geom_point()

```

```{r}
#plot the average rating for each category
dfcat <- subset(df, !is.na(Rating))
dfcat <- dfcat%>%
  group_by(Category)%>%
  summarise(Rating = mean(as.numeric(Rating)))

#plot
ggplot(aes(x = Category, y = Rating), data = dfcat)+
  geom_bar(stat="identity", position="dodge")+
  coord_flip(ylim = c(3.8,4.5))
  
```

>Dating apps have the lowest average rating and family apps have the most 
average rating


# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!